Applications of Technology and Large Data in Statistics Education and Statistical Graphics

### 2015-05-29 DefenseSlides/DefenseSlidesCommitteeHour.Rnw
Further slide edits based on Heike Comments

\begin{frame}
    \frametitle{Constructing Binned Scatterplots}

@@

\begin{frame}
    \frametitle{Rectangular Binning Functions}

@@

where $(\beta_{j-1},\beta_{j}]$ for $j \in \{1,\dots,J\}$ are sequence of 
$J$ adjacent intervals coverting range of the data. (Scott, 1979)

@@

\begin{frame}
    \frametitle{Rectangular Binning Functions}

Standard Rectangular Binning
  \begin{itemize}
    \item Use equally spaced intervals
    \item Bin width $= \omega_X$
    \item $ \{\beta_j \text{ }|\text{ } \beta_j = \beta_{j-1} + \omega_X \}$
    \item $x_j^\ast = (\beta_{j-1} + \beta_j) /2$
  \end{itemize}

@@

\begin{frame}
    \frametitle{Reduced Binned Data}

\tiny
\begin{table}[H]
\begin{minipage}[t]{.33\textwidth}\centering\subfloat[Original Data, 12 rows]{
\begin{tabular}{rr}\hline
       $x$ &  $y$  \\ \hline
-7.7325 & -9.6340 \\ 
  -8.1176 & -1.4529 \\ 
  -5.8996 & -3.2033 \\ 
  -7.0375 & -5.5563 \\ 
  -3.6354 & -3.9315 \\ 
  -8.7639 & 0.9874 \\ 
  -2.9781 & 8.6802 \\ 
  0.8210 & -8.6118 \\ 
  5.4477 & -8.4555 \\ 
  4.6849 & -5.6620 \\ 
  9.4785 & 1.1133 \\ 
  1.7579 & 5.3759 \\  \hline
\end{tabular}}
\end{minipage}\hfil
\begin{minipage}[t]{.33\textwidth}\centering\subfloat[Binned Data Centers, 12 rows]{
\begin{tabularx}{.7\textwidth}{rr} \hline
  $b_X(x)$ & \hspace{.4cm} $b_Y(y)$  \\ \hline
  -5 & -5 \\ 
  -5 & -5 \\ 
  -5 & -5 \\ 
  -5 & -5 \\ 
  -5 & -5 \\ 
  -5 & 5 \\ 
  -5 & 5 \\ 
  5 & -5 \\ 
  5 & -5 \\ 
  5 & -5 \\ 
  5 & 5 \\ 
  5 & 5 \\  \hline
\end{tabularx}}
\end{minipage}\hfil
\begin{minipage}[t]{.33\textwidth}\centering\subfloat[Reduced Binned Data, 4 rows]{
\begin{tabularx}{.7\textwidth}{rrr}\hline
 \hspace{.2cm} $x^\ast$   &   \hspace{.2cm} $y^\ast$ & \hspace{.4cm}c \\ \hline
-5 & -5 & 5 \\ 
  -5 & 5 & 2 \\ 
  5 & -5 & 3 \\ 
  5 & 5 & 2 \\  \hline
\end{tabularx}}
\vfil
\end{minipage}
\caption{\label{DataReductionExample}Original, Binned and Reduced Binned Data Tables, with data storage sizes.  Binned using standard rectangular approach with origin $(\beta_{0,x},\beta_{0,y})$ = (-10,-10) and bin widths $\omega_x = \omega_y = 10$. }
\end{table}
\normalsize
\end{frame}

@@

\begin{frame}
    \frametitle{Rectangular Binning Functions}

Random Rectangular Binning
  \begin{itemize}
    \item Define set of bin centers: $\{x_j^\ast \text{ }|\text{ } x_j^\ast > x_{j-1}^\ast \}$
    \item Assign points at random to surrounding bin centers
    \item Probabilities inversely proportional do distance from bin centers
  \end{itemize}

\begin{eqnarray}\label{randbin1}
b^r_X(x_i) = \left\{\begin{array}{ll} 
  x^\ast_j & \text{ with probability } (x^\ast_{j+1} - x_i)/(x^\ast_{j+1} -x^\ast_{j}) \\
  x^\ast_{j+1} &\text{ with probability }  (x_i - x^\ast_{j})/(x^\ast_{j+1} -x^\ast_{j})
  \end{array}\right\}
\end{eqnarray}

@@

>>>>

\begin{frame}
    \frametitle{Construction of Binned Scatterplots}

@@

\begin{frame}
    \frametitle{Rectangular Binning}

@@

where $(\beta_{j-1},\beta_{j}]$ for $j \in \{1,\dots,J\}$ are sequence of 
$J$ adjacent intervals covering range of the data. (Scott, 1979)

@@

\begin{frame}
    \frametitle{Standard Rectangular Binning}

Standard Rectangular Binning
  \begin{itemize}
    \item Use equally spaced intervals
    \item Bin width $= \omega_X$
    \item Bin origin $= \beta_0 \le x_{(1)}$
    \item $ \{\beta_j \text{ }|\text{ } \beta_j = \beta_{j-1} + \omega_X\}$ where $j \in \{1,\dots,J\}$
    \item $x_j^\ast = (\beta_{j-1} + \beta_j) /2$
  \end{itemize}

@@

\begin{frame}
    \frametitle{Binned Data}

\tiny
\begin{table}[H]
\begin{minipage}[t]{.33\textwidth}\centering\subfloat[Original Data, 12 rows]{
\begin{tabular}{rr}
       $x$ &  $y$  \\
       \hline
-7.7325 & -9.6340 \\ 
  -8.1176 & -1.4529 \\ 
  -5.8996 & -3.2033 \\ 
  -7.0375 & -5.5563 \\ 
  -3.6354 & -3.9315 \\ 
  -8.7639 & 0.9874 \\ 
  -2.9781 & 8.6802 \\ 
  0.8210 & -8.6118 \\ 
  5.4477 & -8.4555 \\ 
  4.6849 & -5.6620 \\ 
  9.4785 & 1.1133 \\ 
  1.7579 & 5.3759 \\ 
\end{tabular}}
\end{minipage}\hfil
\begin{minipage}[t]{.33\textwidth}\centering\subfloat[Binned Data Centers, 12 rows]{
\begin{tabularx}{.7\textwidth}{rr} 
  $b_X(x)$ & \hspace{.4cm} $b_Y(y)$  \\
  \hline
  -5 & -5 \\ 
  -5 & -5 \\ 
  -5 & -5 \\ 
  -5 & -5 \\ 
  -5 & -5 \\ 
  -5 & 5 \\ 
  -5 & 5 \\ 
  5 & -5 \\ 
  5 & -5 \\ 
  5 & -5 \\ 
  5 & 5 \\ 
  5 & 5 \\  
\end{tabularx}}
\end{minipage}\hfil
\begin{minipage}[t]{.33\textwidth}\centering\subfloat[Reduced Binned Data, 4 rows]{
\begin{tabularx}{.9\textwidth}{rrr}

 \hspace{.2cm} $x^\ast$   &   \hspace{.2cm} $y^\ast$ & \hspace{.4cm}c \\ 
 \hline
-5 & -5 & 5 \\ 
  -5 & 5 & 2 \\ 
  5 & -5 & 3 \\ 
  5 & 5 & 2 \\ 
\end{tabularx}}
\vfil
\end{minipage}
\caption{\label{DataReductionExample}Original, Binned and Reduced Binned Data Tables, with data storage sizes.  Binned using standard rectangular approach with binning origin at (-10,-10) and bin widths $\omega_x = \omega_y = 10$. }
\end{table}
\normalsize
\end{frame}

@@

\begin{frame}
    \frametitle{Random Rectangular Binning}

Random Rectangular Binning
  \begin{itemize}
    \item Define set of bin centers: $\{x_j^\ast \text{ }|\text{ } x_j^\ast > x_{j-1}^\ast \}$ for $j \in \{1,\dots,J\}$
    \item $x_1^\ast \le x_{(1)} \le x_{(n)} \le x_J^\ast$
    \item Assign points at random to surrounding bin centers
    \item Probabilities inversely proportional do distance from bin centers
  \end{itemize}

\vspace{.25in}

For $x_i$ such that $x^\ast_{j} \le x_i \le x^\ast_{j+1}$,
\begin{eqnarray}\label{randbin1}
b^r_X(x_i) = \left\{\begin{array}{ll} 
  x^\ast_j & \text{ with probability } (x^\ast_{j+1} - x_i)/(x^\ast_{j+1} -x^\ast_{j}) \\
  x^\ast_{j+1} &\text{ with probability }  (x_i - x^\ast_{j})/(x^\ast_{j+1} -x^\ast_{j})
  \end{array}\right\}
\end{eqnarray}

@@

Cronbach's $\alpha$ measures internal consistency within an item set
%
\begin{equation} \label{eq:alpha}
\alpha = K/(K-1) \cdot \left(1 - \left . \sum_{i=1}^K \V{Y_i} \right /  \V{\sum_{j=1}^K Y_j}\right),
\end{equation}
